---
title: "04_Quiz_In_Class"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Quiz 4 In-Class
Complete this quiz in a .Rmd file. To turn in the quiz, push both a .Rmd file and a knitted .html file to your GitHub site.

Statement of Integrity: Copy and paste the following statement and then sign your name (by typing it) on the line below.

“All work presented is my own. I have not communicated with or worked with anyone else on this quiz.”

Lily Kasperek

Collaboration Reminder: You may not communicate with or work with anyone else on this quiz, but you may use any of our course materials or materials on the Internet.

Question 1 (20 points). Examine the alcohol.csv data set that you may have used as an introductory data set in STAT/DATA 234:

```{r}
library(tidyverse)
library(here)
alcohol_df <- read_csv(here("data/alcohol.csv"))
```

The data set has the following variables on countries throughout the world:

country, the name of the country
beer_servings, the number of beer servings per person per year
spirit_servings, the number of spirit servings per person per year
wine_servings, the number of wine servings per person per year
total_litres_of_pure_alcohol, the number of litres of pure alcohol consumed per person per year

Construct a map of the beer_servings variable. You should choose an appropriate projection and change the default colour scheme of your map.

```{r}
library(maps)
world <- map_data("world")
world

world_full <- left_join(world, alcohol_df, by = c("region" = "country"))
world_full

ggplot(data = world_full, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = beer_servings)) +
  coord_map(projection = "globular", xlim = c(-180, 180)) +
  theme_void() +
  scale_fill_viridis_b()
```

Question 2 (5 points). Give an example of a question of interest about the alcohol data where you would want to construct a visualization other than a map to answer the question. What type of visualization would you construct instead?

One example of a question of interest where you would want to construct a visualization other than a map to answer the question is when mapping the total litres of pure alcohol. One type of visualization to construct instead is a line plot or bar plot. The bar plot could not show all of the countries, but it could be used to show the total liters of alcohol for some of the countries since there is only one observation for each country. 
