---
title: "05_In_Class_Quiz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

__Statement of Integrity:__ Copy and paste the following statement and then sign your name (by typing it) on the line below.

“All work presented is my own. I have not communicated with or worked with anyone else on this quiz.”

Lily Kasperek

__Collaboration Reminder:__ You may not communicate with or work with anyone else on this quiz, but you may use any of our course materials or materials on the Internet.

__Question 1 (7 points).__ Consider the following two bar plots using the palmerpenguins data set. The first is a plot of the penguin species while the second is a plot of the average bill length for each species.

```{r}
library(palmerpenguins)
library(tidyverse)
ggplot(data = penguins, aes(x = species)) +
  geom_bar() +
  labs(y = "Count")

ggplot(data = penguins %>% group_by(species) %>%
         summarise(avg_length = mean(bill_length_mm, na.rm = TRUE)),
       aes(x = species, y = avg_length)) +
  geom_col() +
  labs(y = "Average Bill Length")
```

Which of the two graphs is appropriate to construct? Give a one sentence reason.

Out of the two graphs, the first is appropriate to construct because it is displaying the counts of categorical data. 

__Question 2 (9 points).__ Use the Happy Planet Index data set to construct a graph that does not properly show variability in the underlying data. Recall that some variables in this data set are LifeExpectancy, Wellbeing, Footprint, and Region of the world.

```{r}
library(tidyverse)
hpi_df <- read_csv("~/Desktop/data/hpi-tidy.csv")

hpi_avg <- hpi_df %>% mutate(Region = fct_reorder(.f = Region, .x = Footprint)) %>%
  filter(!is.na(Footprint)) %>% 
  group_by(Region) %>%
  summarise(avg = mean(Footprint))

ggplot(data = hpi_avg, aes(x = Region, y = avg)) +
  geom_col() +
  coord_flip() +
  labs(y = "Average Footprint")
```

Question 3 (7 points). Fix your graph from the previous question so that it does properly show variability in the underlying data.

```{r}
hpi_fix <- hpi_df %>% mutate(Region = fct_reorder(.f = Region, .x = Footprint)) %>%
  filter(!is.na(Footprint))

ggplot(data = hpi_fix, aes(x = Region, y = Footprint)) +
  geom_point(alpha = 0.4) +
  geom_point(data = hpi_avg, aes(y = avg, colour = "Mean Footprint")) +
  coord_flip() +
   scale_colour_manual(values = c("Mean Footprint" = "red")) +
  labs(colour = " ")
```

The remaining 2 points are given for committing and pushing your .Rmd and .html files correctly to GitHub.